Code Club Project: Assessing whether intra and inter-genomic variation hinder utility of ESVs and ASVs

### 2021-01-15 submission/manuscript.Rmd
Replicate threshold to get one OTU per genome/species 100 times

A method to avoid splitting a single genome into multiple units of inference is to cluster 16S rRNA gene sequences together that are similar. However, this also increases the risk of lumping together genes from different species that are similar to each other. Therefore, I assessed the impact of the threshold used to define clusters of 16S rRNA genes on the propopensity to lump species together and split genome apart. I identified the threshold where 90% of bacterial species would be represented by a single OTU. For full length 16S rRNA gene sequences, I found that at a threshold of XX%, 90% of the species would be represented by a single OTU. Similarly, thresholds of XX, XX, and XX% were observed for the V4, V3-V4, and V4-V5 regions. However, at these thresholds, multiple species could be represented by the same OTU. At the highest level of resolution, XX% of the species shared a 16S rRNA gene sequence variant with another species. Given the risk of splitting a genome into multiple OTUs is more biologically problematic than lumping species together, larger thresholds are advisable.

>>>>

```{r}
thresholds <- read_tsv(here("data/processed/thresholds_for_single_otu.tsv"),
											 col_types=cols(region = col_character(),
											 							 .default = col_double())) %>%
	filter(n_rrns == 7)

threshold_v19 <- thresholds %>% filter(region == "v19") %>% pull(threshold) * 100
threshold_v4 <- thresholds %>% filter(region == "v4") %>% pull(threshold) * 100
threshold_v34 <- thresholds %>% filter(region == "v34") %>% pull(threshold) * 100
threshold_v45 <- thresholds %>% filter(region == "v45") %>% pull(threshold) * 100
```

A method to avoid splitting a single genome into multiple units of inference is to cluster 16S rRNA gene sequences together that are similar. However, this also increases the risk of lumping together genes from different species that are similar to each other. Therefore, I assessed the impact of the threshold used to define clusters of 16S rRNA genes on the propensity to split a genome apart or to lump species together. For full length 16S rRNA gene sequences, I found that at a threshold of `r format_pct(threshold_v19)`%, 95% of the species with 7 copies of the rrn operon would be represented by a single OTU. Similarly, thresholds of `r format_pct(threshold_v4)`, `r format_pct(threshold_v34)`, and `r format_pct(threshold_v45)`% were observed for the V4, V3-V4, and V4-V5 regions, respectively. However, at these thresholds, multiple species could be represented by the same OTU. At the highest level of resolution, XX% of the species shared a 16S rRNA gene sequence variant with another species. Given the risk of splitting a genome into multiple OTUs is more biologically problematic than lumping species together, larger thresholds are advisable.

### 2021-01-15 submission/manuscript.tex
Replicate threshold to get one OTU per genome/species 100 times

A method to avoid splitting a single genome into multiple units of
inference is to cluster 16S rRNA gene sequences together that are
similar. However, this also increases the risk of lumping together genes
from different species that are similar to each other. Therefore, I
assessed the impact of the threshold used to define clusters of 16S rRNA
genes on the propopensity to lump species together and split genome
apart. I identified the threshold where 90\% of bacterial species would
be represented by a single OTU. For full length 16S rRNA gene sequences,
I found that at a threshold of XX\%, 90\% of the species would be
represented by a single OTU. Similarly, thresholds of XX, XX, and XX\%
were observed for the V4, V3-V4, and V4-V5 regions. However, at these
thresholds, multiple species could be represented by the same OTU. At
the highest level of resolution, XX\% of the species shared a 16S rRNA
gene sequence variant with another species. Given the risk of splitting
a genome into multiple OTUs is more biologically problematic than
lumping species together, larger thresholds are advisable.

>>>>

A method to avoid splitting a single genome into multiple units of
inference is to cluster 16S rRNA gene sequences together that are
similar. However, this also increases the risk of lumping together genes
from different species that are similar to each other. Therefore, I
assessed the impact of the threshold used to define clusters of 16S rRNA
genes on the propensity to split a genome apart or to lump species
together. For full length 16S rRNA gene sequences, I found that at a
threshold of \textbf{5.5}\%, 95\% of the species with 7 copies of the
rrn operon would be represented by a single OTU. Similarly, thresholds
of \textbf{2.5}, \textbf{4.0}, and \textbf{3.5}\% were observed for the
V4, V3-V4, and V4-V5 regions, respectively. However, at these
thresholds, multiple species could be represented by the same OTU. At
the highest level of resolution, XX\% of the species shared a 16S rRNA
gene sequence variant with another species. Given the risk of splitting
a genome into multiple OTUs is more biologically problematic than
lumping species together, larger thresholds are advisable.
